Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits

### 2022-10-20 04_paper/10_20_draft_newabstract.Rmd
Changes
Update abstract, add placeholders for variables explanations, and polish section 6 stylized facts

abstract: "Absent formal avenues of transparency, how do China’s foreign aid and debt allocation relate to public diplomacy and international recognition efforts? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. These measures capture two distinct but related aspects of China’s foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. We also distinguish Beijing-as-donor and Beijing-as-lender. We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. To the extent that China’s development finance complements these goals, we expect greater coverage of a developing country in a Chinese media outlet directed to foreign readers and greater bilateral diplomatic activity to correlate with greater Chinese foreign aid and debt giving. To test this expectation, we leverage (1) AidData’s Chinese development finance datasets and yearly counts of bilateral official diplomatic visits from China and (2) the meta-data of millions of English edition Xinhua news articles from 2000 to 2017 in the Cline Center for Advanced Social Research. We find that China pays more media attention to aid recipients but engages more in diplomatic visits with debtors. Aid recipient countries disproportionately get more media attention from Xinhua via news stories that report both the state of their economy and their developmental needs. Notably, Xinhua does not just promote China’s efforts but also gives a round picture of the countries’ economic conditions in justifying their legitimate need for aid. Our results dispel the myths of China as a ’rouge donor’ and Chinese ’debt-trap diplomacy.’ We suggest that much like traditional donors, China’s foreign aid presents a versatile foreign policy tool through which Beijing seeks to accomplish important objectives in international politics. Also, China is more likely to provide loans to diplomatically aligned countries."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

@@

On the one hand, since aid is supposed to have more direct links to China's foreign policy goals, it follows that China will complement its aid giving with visible acts of diplomacy and efforts to increase the salience of aid recipients in foreign audiences. On the other hand, while aid and debt may have different goals, such a tidy distinction may be too simplistic. Perhaps more importantly, some of the market oriented objectives that drive China's global debt might benefit from public displays of diplomacy. On this front, recent research suggests that Beijing uses its debt to ensure access to natural resources that China lacks domestically. As noted above, @dreherEtAl2021 find that Beijing targets more of its loans in resource rich developing countries, a strategy that @dreher_banking_2022 contend is meant to guarantee access to said resources. When Beijing's policy banks provide these loans, developing country governments collateralize their oil or rare minerals, all but assuring access to these resources for Beijing since many of these countries are unable to directly repay the loans[@dreher_banking_2022].

@@

Figure \@ref(fig:total-aid-debt) shows total ODA (aid) and OOF (debt) disbursed by Beijing per year from 2002 to 2017. Values are shown in millions of 2017 USD. Blue shows values for aid and red shows values for debt. The difference in China's application of debt relative to aid is stark, and the gulf only expands over time.

This explosion in debt relative to aid aligns with a shift in Beijing's strategy going into the twenty-first century. Burdened with vast foreign currency reserves due to years-long trade surpluses, confronted with excess domestic production of industrial goods, and in need of access to resources to support its continued economic growth, Beijing tasked its so-called policy banks to help Chinese firms gain access to foreign markets [@dreher_banking_2022]. This focus spurred a radical change in China's development finance portfolio. As @dreher_banking_2022 put it, "Beijing began to behave less like a benefactor and more like a banker" (5). Figure \@ref(fig:total-aid-debt) shows this transformation from donor to lender in the starkest of terms.

```{r total-aid-debt, fig.cap='Total yearly ODA (aid) and OOF (debt) issued by China from 2002 to 2017 in millions of 2017 constant USD.'}
library(knitr)
library(here)
include_graphics(
  here('06_figures/aid_debt_trends.png')
)
```

```{r 4-types, fig.cap='Number of developing countries per year that are (a) beneficiaries of Chinese aid, (b) debtors, (c) both, and (d) neither.'}
include_graphics(
  here('06_figures/dev_countries_over_time_by_type.png')
)
```

@@

Figure \@ref(fig:aid-debt-country) shows the top-20 Chinese aid and debts recipient countries on average between 2002 and 2017. Notably, on average, Iraq received 500 million in aid but borrows 0 dollars in loans from China. In contrast, Russia borrowed almost 8 billion dollars from China but received no aid at all. There is little overlap between the top recipient countries that receive both aid and debt from China. There are stark variations in terms of the amounts China lends. Russia and Venezuela were the only countries that borrowed more than 5 billion dollars in debt on average, the sum of which accounted for about 30% of the total lending amounts of the top-20 debt recipients. Eight countries are shown on both the left and right frames: Indonesia, Pakistan, Laos, Ethiopia, Cambodia, Sri Lanka, Kenya, and Sudan. Geographically, five are in South or East Asia, whereas three are in Africa.

@@

@@

Recall that hypothesis 1 holds that Beijing will concentrate more *Xinhua* coverage in development finance recipients, and that hypothesis 2 holds that it will send more diplomatic missions to recipients. With respect to aid, there is little observable difference in country coverage in *Xinhua* depending on recipient status. However, recipients host appreciably more diplomatic missions from Beijing than non-recipients. The difference narrows by 2013, but remains across the entire sample period. With respect to debt, there are notable differences in both *Xinhua* mentions and diplomatic missions between debtor and non-debtor developing countries. Debtor status corresponds with more frequent yearly mentions in *Xinhua* and a greater number of hosted missions from Beijing.

These patterns remain only suggestive. Many factors jointly determine Beijing's financing, *Xinhua* coverage, and diplomatic missions that likely confound straightforward identification of the relationships among these variables. The next section summarizes the results from more rigorous statistical tests of hypotheses 1 and 2.

@@

```{r binom-est, fig.cap='Estimated Negative Binomial model coefficients with 95\\% confidence intervals. Controls and year fixed effects not shown. Models were estimated with random recipient intercepts.'}
include_graphics(
  here('06_figures/neg_binom_estimates.png')
)
```

```{r predicted-diff, fig.cap='Counterfactual estimates of yearly \\textit{Xinhua} mentions and diplomatic missions from Beijing.'}
include_graphics(
  here('06_figures/predicted_differences.png')
)
```

>>>>

abstract: "Absent formal avenues of transparency, how do China’s foreign aid and debt allocation relate to public diplomacy and international recognition efforts? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. We expect greater coverage of a developing country in a Chinese media outlet directed to aid giving, and greater bilateral diplomatic activity to correlate with greater debt allocation. To test this expectation, we leverage (1) AidData’s Chinese development finance datasets and (2) the meta-data of millions of Xinhua news articles in the Cline Center for Advanced Social Research. We find that China pays more media attention to aid recipients but engages more in diplomatic visits with debtors. Aid-recipient countries disproportionately get more media coverage on the state of their economy and developmental needs. Also, China is more likely to provide loans to diplomatically aligned countries. Our results dispel the myths of China as a ’rouge donor’ and Chinese ’debt-trap diplomacy.’ We suggest that much like traditional donors, China’s foreign aid presents a versatile foreign policy tool in international politics. "
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```
\clearpage

@@

On the one hand, since aid is supposed to have more direct links to China's foreign policy goals, it follows that China will complement its aid giving with visible acts of diplomacy and efforts to increase the salience of aid recipients in foreign audiences. On the other hand, while aid and debt may have distinct goals, such a tidy distinction may be too simplistic. Perhaps more importantly, some of the market oriented objectives that drive China's global debt might benefit from public displays of diplomacy. On this front, recent research suggests that Beijing uses its debt to ensure access to natural resources that China lacks domestically. As noted above, @dreherEtAl2021 find that Beijing targets more of its loans in resource rich developing countries, a strategy that @dreher_banking_2022 contend is meant to guarantee access to said resources. When Beijing's policy banks provide these loans, developing country governments collateralize their oil or rare minerals, all but assuring access to these resources for Beijing since many of these countries are unable to directly repay the loans[@dreher_banking_2022].

@@

In this section, we will provide some figures to show the trends of Chinese aid and debt between 2002 and 2017 and the top 20 recipients in terms of their financial portfolios on average. Then we will present some visuals on the correlations between *Xinhua* news coverage, diplomatic visits, Chinese financial outflow destinations in yearly trends as well as in countries ranked the highest percentiles on each dimension.

## Main variables

*aid, debt, Xinhua mentions, diplomatic visits*

## Trends in Aid and Debt

Figure \@ref(fig:total-aid-debt) shows total ODA (aid) and OOF (debt) disbursed by Beijing per year from 2002 to 2017. Values are shown in millions of 2017 USD. Blue shows values for aid and red shows values for debt. The difference in China's application of debt relative to aid is stark, and the gulf only expands over time. This explosion in debt relative to aid aligns with a shift in Beijing's strategy going into the twenty-first century. Burdened with vast foreign currency reserves due to years-long trade surpluses, confronted with excess domestic production of industrial goods, and in need of access to resources to support its continued economic growth, Beijing tasked its so-called policy banks to help Chinese firms gain access to foreign markets [@dreher_banking_2022]. This focus spurred a radical change in China's development finance portfolio. As @dreher_banking_2022 put it, "Beijing began to behave less like a benefactor and more like a banker" (p. 5). Figure \@ref(fig:total-aid-debt) shows this transformation from donor to lender in the starkest of terms.

```{r total-aid-debt, fig.cap='Total yearly ODA (aid) and OOF (debt) issued by China from 2002 to 2017 in millions of 2017 constant USD.', out.width="80%", fig.align = 'center'}
library(knitr)
library(here)
include_graphics(
  here('06_figures/aid_debt_trends.png')
)
```

```{r 4-types, fig.cap='Number of developing countries per year that are (a) beneficiaries of Chinese aid, (b) debtors, (c) both, and (d) neither.', out.width="80%", fig.align = 'center'}
include_graphics(
  here('06_figures/dev_countries_over_time_by_type.png')
)
```

@@

Figure \@ref(fig:aid-debt-country) shows the top 20 Chinese aid and debts recipient countries on average between 2002 and 2017. Notably, on average, Iraq received 500 million in aid but borrows 0 dollars in loans from China. In contrast, Russia borrowed almost 8 billion dollars from China but received no aid at all. There is little overlap between the top recipient countries that receive both aid and debt from China. There are stark variations in terms of the amounts China lends. Russia and Venezuela were the only countries that borrowed more than 5 billion dollars in debt on average, the sum of which accounted for about 30% of the total lending amounts of the top-20 debt recipients. Eight countries are shown on both the left and right frames: Indonesia, Pakistan, Laos, Ethiopia, Cambodia, Sri Lanka, Kenya, and Sudan. Geographically, five are in South or East Asia, whereas three are in Africa.

@@

## Interplay between Chinese Aid and Debt Outflows, News Mentions and Diplomatic Destinations

@@

Recall that hypotheses 1 and 3 hold that Beijing will concentrate more *Xinhua* coverage in development finance recipients, and that hypotheses 2 and 4 hold that it will send more diplomatic missions to recipients. With respect to aid, there is little observable difference in country coverage in *Xinhua* depending on recipient status. However, recipients host appreciably more diplomatic missions from Beijing than non-recipients. The difference narrows by 2013, but remains across the entire sample period. With respect to debt, there are notable differences in both *Xinhua* mentions and diplomatic missions between debtor and non-debtor developing countries. Debtor status corresponds with more frequent yearly mentions in *Xinhua* and a greater number of hosted missions from Beijing.

Finally, we examine the overlaps between top 20 *Xinhua* news coverage countries and the top 10 percent aid and debt recipients of the year in Figure \@ref(fig:coverage-combined). The y-axis is the list of top 20 *Xinhuan* news coverage countries ranked in order. The blud or red colors denote whether the countries are ranked as top 10 percent debt aid or recipients of the year. The bigger the dot is, the higher amount of yearly news coverage count it is. Russia received the highest number of news exposure between 2002 and 2017 on average. It also consistently stands as top 10 percent debt recipients of the year (shown as red dots in Figure \@ref(fig:coverage-combined) (bottom)). Take India as another example: in 2009, 2010, 2011 and 2012, it received over 400 times mentions in *Xinhua* and ranked as the top 10 percent debt recipients. Indonesia also consistently received over 200 yearly mentions in *Xinhua* reports and borrowed significantly amount of money from China. Overall, there is a significant overlap for countries receiving high media exposure and scoring high percentiles in Chinese debt and aid during the time of investigation.

```{r coverage-combined, fig.cap='Distribution of top 20 Xinhua news coverage ocuntries and top 10 percent aid (top) and debt (bottom) recipients from 2002 to 2017.', out.width="100%",fig.show='hold'}

include_graphics(
  here('06_figures/coverages_combined.jpg'))

```

Similarly, Figure \@ref(fig:visits-combined) shows the overlaps between top 20 diplomatic visits countries and the top 10 percent aid and debt recipients of the year. Nine countries[^7] were listed as the top 20 diplomatic visits and *Xinhua* news coverage on average between 2002 and 2017. Countries such as Kazakhstan, India and Thailand often hosted Chinese high officials for more than 30 times a year. Overall, China seemed to visit the list of the top 10 countries over 10 times and reguarly except for four countries in 2016 and 2017. These countries were more likely to be countries where received the highest percentiles of financial inflows from China.

[^7]: India, Indonesia, Thailand, Malaysia, Afghanistan, Pakistan, Philippines, Cambodia, Vietnam.

```{r visits-combined, fig.cap='Distribution of top 20 diplomatic visits destinations and top 10 percent aid (top) and debt (bottom) recipients from 2002 to 2017.', out.width="100%", fig.show='hold'}

include_graphics(
  here('06_figures/visits_combined.jpg'))

```

These patterns remain only suggestive. Many factors jointly determine Beijing's financing, *Xinhua* coverage, and diplomatic missions that likely confound straightforward identification of the relationships among these variables. The next section summarizes the results from more rigorous statistical tests of hypotheses 1 to 4.

@@

**Adding confounders here**

```{r binom-est, fig.cap='Estimated Negative Binomial model coefficients with 95\\% confidence intervals. Controls and year fixed effects not shown. Models were estimated with random recipient intercepts.', fig.show='hold'}
include_graphics(
  here('06_figures/neg_binom_estimates.png')
)
```

```{r predicted-diff, fig.cap='Counterfactual estimates of yearly \\textit{Xinhua} mentions and diplomatic missions from Beijing.', fig.show='hold'}
include_graphics(
  here('06_figures/predicted_differences.png')
)
```
